Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow

### 2014-07-28 writeups/dissertation/dissertation.tex
foo

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
  The task of choosing a hashtag for a tweet and tagging a post on StackOverflow was framed as a declarative memory retrieval problem.
  Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random permutation vector-based model.
  The results show that past user behavior of tag use is a strong predictor of future performance, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, the common natural-language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
  and an entropy-weighting technique was developed that improved model accuracy by more efficiently allocating limited attentional resources to the predictive words in the posts and tweets.
  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order,
  which shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human declarative memory.
The declarative memory component in ACT-R is a Bayesian statistical model.
This is the same type of co-occurrence Naive Bayes statistical model that is prominent in the more general machine-learning field.
When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
The primary reason for this is that the stability of statistical models increase as sample size increases.

@@

This approach is problematic for two main reasons:
First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
Second, it means that modelers are continuing to use small sample sizes for the declarative memory components in their models.

@@

A growing number of users on social media sites are focusing on and looking for recently created information when performing a search \parencite{Jansen2011}.
In other words, users on social media sites want the fresh \emph{stream} of information about a particular topic.
This can certainly be seen in the main user-interface views that these social media sites provide:
Twitter's homepage showing followees' tweets, Facebook's homepage showing friends' posts, and StackOverflow's daily digest of posts for particular tags.

@@

Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010}.
Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
Two of the current state-of-the-art memory retrieval models are based on a co-occurrence methodology, and I will be comparing these two models for hashtag prediction:
ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.

@@

I will compare these two memory systems across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#''.
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

@@

These memory retrieval models will also be tested on tag use for StackOverflow posts.
StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
User-chosen tags on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.

@@

ACT-R's declarative memory and vector-based memory systems are substantially different in their mathematical structure.
Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., \cite{Rutledge2008} modeling the ``fan effect'').
I will extend the work done by \citeauthor{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
I will also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random permutation model \parencite{Sahlgren2008}.

An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
Word order information on sites like Twitter may contain highly predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.
I will explore how the word-order strength of vector-based models can be incorporated into the ACT-R theory when testing the models on hashtag prediction tasks.

ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2007}.
This decay rate equation formalizes how recency and frequency of hashtag use relate to the prior probability that a hashtag will be retrieved.
However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.
I will explore how a user's prior knowledge influences the likelihood that a particular hashtag is chosen for both ACT-R's memory theory and vector-based models.

@@

Prior research has examined the growth and decay cycle of hashtag use across users \parencite{Tsur2012}.
However, much less is known about hashtag lifetime for individual users.
It may not necessarily follow that a hashtag's life cycle for individual users matches the life cycle across users.
Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into a specific user's prior likelihood of hashtag use.
So I will characterize the hashtag growth and decay cycle for individual users.
The expectation is that this cycle will match ACT-R's decay rate used in the declarative memory system to compute a particular memory item's prior.

@@

Base-level activation reflects the log prior odds of needing an observed chunk again.
The primary way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its current frequency and recency of use.

@@

The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
It is important that this measure scales well enough to handle the sheer volume of information present in online social media datasets.
Large-scale social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large-scale human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

@@

Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.

@@

ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
This type of information on a user's prior behavior is not being utilized with vector-based models currently.
It may certainly be the case that vector-based models can be modified to incorporate base-level activation, and this idea will be pursued with this research.

@@

Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
These two computations are certainly not equivalent, even though they are most likely correlated with one another.
It may be interesting to explore how model performance changes when one calculation is swapped out for another.
For example, a vector-based model could generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
It is unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.

@@

The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
Netflix is an online subscription service where users can stream their favorite television episodes and movies.
Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
In order to provide a high quality user experience on the site, Netflix created a content-recommendation service.
This tool recommends to the user movies and television episodes that most likely match their specific interests.

@@

Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
Google has even deployed a hashtag-recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.

@@

It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular environment
(i.e., when predicting the user-chosen tags for human-created content on social media sites).

@@

\textcite{Stanley2013} used an ACT-R inspired Bayesian memory model to predict human tagging on the StackOverflow site, but to our knowledge no vector-based memory model has been tested or used.
Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}. 
Given that the Bayesian cognitively-plausible models have done so well to predict tag use on StackOverflow, it will be interesting to see how a cognitively-plausible vector-based model performs at the task.
Prediction differences between the two models may help reveal ways where the constraints of each model can be adjusted in order to improve accuracy.

Also, a vector-based retrieval model has not yet been tested on Twitter.
One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
It will be interesting to see how much (if at all) accuracy improves when incorporating word order information into the model.

@@

This preliminary research is encouraging, and suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
It also seems plausible that model accuracy can improve an even greater amount by
[1] figuring out the optimal way to combine global tag history with a user's past tagging history, and
[2] using the time-dependent form of the base-level component in Table \ref{tabACTRModel}, in order to account for the fact that a user's tagging preferences change over time.
So these proposed modifications will be explored further when incorporating a user's past tagging history into the models for this research.

@@

Two models will be evaluated for hashtag prediction on both the Twitter and StackOverflow datasets:
A vector-based model and Bayesian retrieval model.
The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} will be used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel} will be used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.

These models will be the starting point for modeling tag use on the Twitter and StackOverflow sites.
The models may be modified during testing if the data indicate that modifications will increase the accuracy in which the model's recommended tags agree with the post author's chosen tags.
Also there are planned modifications to each model as part of the research, as these modifications seem likely to improve performance: 
incorporating user prior information into the vector-based model and incorporating word order information into the Bayesian model.

@@

The two memory models will be evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
Parameters for each model will be calibrated on a large subset of the Twitter and StackOverflow datasets.
Each model will be tested by evaluating the accuracy of model-chosen tags on a fresh test subset of the datasets.
Accuracy will be assessed by comparing model performance with the human data.

% Started re-reading and re-writing here.

@@

A popular-users dataset was collected for both StackOverflow and Twitter.
This dataset consists of all of the tweets from 24 different samples of the top users for each site.
This dataset was collected by fetching all tweet history for each of the top users.
It was used to explore the growth and decay characteristics of a hashtag for individual users.
The temporal dynamics of ACT-R's decay rate model for a chunk were evaluated against this data.

@@

Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
It took around four days to collect a single dataset, and one collection was done each week.
Multiple datasets were collected to ensure that the results weren't dependent on a particular sample of hashtags, and would generalize across different popular-hashtag datasets and different time periods.

@@

The entire StackOverflow dataset is packaged and released quarterly to the public.
So this package was downloaded and the database tables were imported into PostgreSQL.

@@

A custom tokenization algorithm created by \textcite{Owoputi2013} was used to chunk the words in the text.
This algorithm was specifically created for tokenizing Twitter content.
However during testing it was found to be robust and well suited for StackOverflow content.
The initial model results with the new tokenizer were qualitatively very similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
This is most likely because the language for the samples taken for both sites is informal English,
and it seams reasonable to think that tokenizing all of the informal slang, jargon, and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar for both algorithms on StackOverflow,
and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.

@@

\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing 'ed' and 'ing' endings), but
with the StackOverflow dataset, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
This can be thought of as semantic stemming, where two words with identical meaning are converted to the same canonical word.
So tag words in the post and used tags for the post were converted to their root tag when possible.

@@

``Stop words'' are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
It is typically recommended \parencite{Bird2009} to remove stop words such as ``the'', ``a'', ``or'' when building the co-occurrence matrix.
Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in text.
Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.

Stop words are commonly removed by either filtering based on a pre-determined list or filtering based on the high-frequency words that occur in the dataset.
\textcite{Sahlgren2008} tried both methods to remove stop words for the random permutation vector-based model, and found that the data-driven approach
(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a pre-determined list.

@@

All three stop word techniques (i.e., pre-determined list, high-frequency filtering, and entropy weighting) were explored for this analysis.

@@

The decay rate parameter $d$ represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
Higher values mean that chunk activation decays faster, which means that the most recently-presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occurred most frequently will have the highest activation,
and activation depends less on the specific chunk that have been observed in the recent past.

@@

The full StackOverflow dataset and Twitter popular-users dataset were used to characterize tag growth and decay for a user on each site.
Each user's tag behavior from the popular-users dataset for each site was analyzed.

@@

20 different values of the decay rate parameter were explored for each slice
(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
The values were chosen in order to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d=.5$), to a pure recency model ($d=20$).
The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
However, only values equal to or below 1 were examined for the optimized learning form since values above 1 for this form cause a bifurcation and are not psychologically interesting.

Both the standard form and optimized learning form of $B_{i}$ were used to predict each user's specific tag history.
To do this, a user's tag history was collected, and then each tag instance from oldest to newest was examined.
For each of these tagging instances, the model generated an activation value for each tag based on that tag's frequency and recency of use. 
This means, for example, that the model always missed characterizing the first observed tag for each user, since at that point the model does not have any history of tag use for that user.
This also means that the tag space for each user is constrained and limited to the tags that each user has used in the past.

@@

The specific value of the hashtag used for both plots is nominal and not important for this analysis.
What is interesting is how large the tagging space is for these users.
There are bands of commonly-used tags for each user, but there are also instances where new tags are used, and a large percentage of tags used are not in the most common bands.
This helps illustrate how challenging modeling this task can be.
The model's goal is to predict at each instance what tag will be chosen, but since the tag space is so large and varied, this task is quite challenging.

Although the tag space for the Twitter user is smaller than the StackOverflow user in these plots, this is not a consistent trend.
These plots were chosen from the set to also show that the amount of tag variation for individual users also varies: some use many tags and some use just a few.
To accurately model human performance in this domain, the model must be able to make accurate predictions across all variations of user tagging behavior.

@@

>>>>

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
  The task of choosing a hashtag for a tweet and tagging a post on StackOverflow was framed as a declarative memory retrieval problem.
  Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random permutation vector-based model.
  The results show that past user behavior of tag use is a strong predictor of future behavior, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, the common natural language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
  and an entropy-weighting technique was developed that improved model accuracy by more efficiently allocating limited attentional resources to the predictive words in the posts and tweets.
  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order,
  which shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human declarative memory.
The declarative memory component in ACT-R is a Bayesian statistical model.
This is the same type of co-occurrence Naive Bayes statistical model that is prominent in the more general machine learning field.
When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
The primary reason for this is that the stability of statistical models increase as sample size increases.

@@

But this approach is problematic for two main reasons:
First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
Second, it means that modelers are continuing to use small sample sizes for the declarative memory components in their models.

@@

A growing number of users on social media sites are focusing on and looking for recently created information when performing a search \parencite{Jansen2011}.
In other words, users on social media sites want the fresh \emph{stream} of information about a particular topic.
This can certainly be seen in the main user interface views that these social media sites provide:
Twitter's homepage showing followees' tweets, Facebook's homepage showing friends' posts, and StackOverflow's daily digest of posts for particular tags.

@@

Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010}.
Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
At least two of the current memory retrieval models are based on a co-occurrence methodology, and I will be comparing these two models for hashtag prediction:
ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.

@@

These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#''.
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

@@

These memory retrieval models were also tested on tag use for StackOverflow posts.
StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
The tags chosen by a user on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.

@@

ACT-R's declarative memory and vector-based memory systems are substantially different in their mathematical structure.
Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., \cite{Rutledge2008} modeling the ``fan effect'').

An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
Word order information on sites like Twitter may contain highly predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.

ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2007}.
This decay rate equation formalizes how recency and frequency of hashtag use relate to the prior probability that a hashtag will be retrieved.
However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.

@@

Prior research has examined the growth and decay cycle of hashtag use across users \parencite{Tsur2012}.
However, much less is known about hashtag lifetime for individual users.
It may not necessarily follow that a hashtag's life cycle for individual users matches the life cycle across users.
Further, modeling a hashtag's life within users is much more applicable to hashtag prediction, since that model can be directly incorporated into a specific user's prior likelihood of hashtag use.

@@

Base-level activation reflects the log prior odds of needing an observed chunk again.
The default way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its current frequency and recency of use.

@@

The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
It is important that this measure scales well enough to handle the sheer volume of information present in online social media datasets.
Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

@@

Another issue with LSA is that adding additional observations to an previously-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
Since computing the SVD is the most computationally expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.

@@

ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
This type of information on a user's prior behavior is not being utilized with vector-based models currently.

@@

Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
These two computations are certainly not equivalent, even though they are most likely correlated with one another.

@@

The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
Netflix is an online subscription service where users can stream their favorite television episodes and movies.
Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
In order to provide a high quality user experience on the site, Netflix created a content recommendation service.
This tool recommends to the user movies and television episodes that most likely match their specific interests.

@@

Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.

@@

It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
This suggests that more cognitively-plausible memory retrieval models may outperform more general machine learning approaches in this particular environment
(i.e., when predicting the user-chosen tags for human-created content on social media sites).

@@

\textcite{Stanley2013} used an ACT-R inspired Bayesian memory model to predict human tagging on the StackOverflow site, but to our knowledge no vector-based memory model has been tested or used.
Both types of models are neurologically plausible, and both have been successfully incorporated as the declarative memory component of the ACT-R cognitive architecture \parencite{Rutledge2007}.

Also, a vector-based retrieval model has not yet been tested on Twitter.
One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.

@@

This preliminary research is encouraging, and suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
It also seems plausible that model accuracy can improve an even greater amount by using the time-dependent form of the base-level component in Table \ref{tabACTRModel},
in order to account for the fact that a user's tagging preferences change over time.

@@

Two models were evaluated for hashtag prediction on both the Twitter and StackOverflow datasets:
A vector-based model and Bayesian retrieval model.
The random permutation model \parencite{Sahlgren2008} in Table \ref{tabRandPermModel} was used for the vector-based model, since it has shown to perform better than the BEAGLE model \parencite{Recchia2010}.
The modified ACT-R Bayesian model \parencite{Stanley2013} in Table \ref{tabModACTRModel} was used for the Bayesian model, since it has already been applied to tag use on StackOverflow posts.

@@

The two memory models were evaluated on predicting user-chosen tags for Twitter tweets and StackOverflow posts.
The co-occurrence matrices used for the context component of each model were generated from a large subset of the Twitter and StackOverflow datasets.
Each model was tested by evaluating the accuracy of model-chosen tags on a fresh test subset of the datasets.
Accuracy was assessed by comparing model performance with the human data.

% Started rereading and rewriting here.

@@

A popular-users dataset was collected for both StackOverflow and Twitter.
This dataset consists of all of the tweets from various samples of the top users for each site.
This dataset was collected by fetching all past tagging history for each of the top users.
It was used to explore the growth and decay characteristics of a hashtag for individual users.
The temporal dynamics of ACT-R's decay rate model for a chunk were evaluated against this data.

@@

Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
Multiple datasets were collected to ensure that the results weren't dependent on a particular sample of hashtags, and would generalize across different popular-hashtag datasets and different time periods.
Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.

@@

The entire StackOverflow dataset is packaged and released quarterly to the public.
So the packaged dataset was downloaded and the database tables were imported into PostgreSQL.

@@

A custom tokenization algorithm created by \textcite{Owoputi2013} was used to chunk the words in the text.
This algorithm was specifically created for tokenizing Twitter content.
However during testing it was found to be robust and well suited for StackOverflow content.
The initial model results with the new tokenizer were qualitatively very similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
This is most likely because the language for the samples taken for both sites is informal English,
and it seams reasonable to think that tokenizing all of the informal language and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar for both algorithms on StackOverflow,
and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.

@@

\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing 'ed' and 'ing' endings), but
with the StackOverflow dataset, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
This can be thought of as semantic stemming, where two words with identical meaning are converted to the same canonical word.
So tags used for the post were converted to their root tag when possible.

@@

``Stop words'' (e.g., ``the'', ``a'', ``or'') are commonly used words that co-occur with all tags (i.e., have little to no discriminating power).
It is typically recommended \parencite{Bird2009} to remove stop words when building the co-occurrence matrix.
Otherwise the matrix may not scale as well since these low-predictor stop words commonly appear in text.
Also for vector-based systems, memory vectors will become overly saturated with the environment vectors of the commonly-occurring stop words.

Stop words are commonly removed by either filtering based on a predetermined list or filtering based on the high-frequency words that occur in the dataset.
\textcite{Sahlgren2008} tried both methods to remove stop words for the random permutation vector-based model, and found that the data-driven approach
(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a predetermined list.

@@

All three stop word techniques (i.e., predetermined list, high-frequency filtering, and entropy weighting) were explored for this analysis.

@@

The decay rate parameter $d$ represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
Higher values mean that chunk activation decays faster, which means that the most recently presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occurred most frequently will have the highest activation,
and activation depends less on the specific chunk that have been observed in the recent past.

@@

The StackOverflow and Twitter popular-users datasets were used to characterize tag growth and decay for a user on each site.
Each user's tag behavior from the popular-users dataset for each site was analyzed.

@@

20 different values of the decay rate parameter were explored for each slice
(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
The values were chosen to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d=.5$), to a pure recency model ($d=20$).
The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
However, only values equal to or below 1 were examined for the optimized learning form since values above 1 for this form cause a bifurcation and are not psychologically interesting.

Both the standard form and optimized learning form of $B_{i}$ were used to predict each user's specific tag history.
To do this, a user's entire tag history was collected, and then each tag instance from oldest to newest was examined.
For each of these tagging instances, the model generated an activation value for each tag based on that tag's frequency and recency of use. 
This means, for example, that the model always missed characterizing the first observed tag for each user, since at that point the model does not have any history of tag use for that user.
This also means that the tag space for each user is constrained and limited to the tags that each user has used in the past.

@@

The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
What is interesting is the large tag space for these users.
There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
This helps illustrate how challenging modeling this task can be.
The model's goal is to predict at each instance what tag will be chosen, but since the tag space is so large and varied, the model must predict the correct few tags that are used from a much larger set of possible tags.

Although the tag space for the Twitter user is smaller than the StackOverflow user in these plots, this is not a consistent trend.
These plots were chosen from the set to also show that the amount of tag variation for individual users also varies: some use many tags and some use just a few.
To accurately model human performance in this domain, the model must be able to make correct predictions across all variations of user tagging behavior.

@@

That said, comparing the effect size between changing the decay rate value and using the standard form,
model accuracy improves much more when the standard form replaces the optimized learning form than when the decay rate value is increased from 0.5 to 0.8.
For these datasets, it is much more important to use the standard form of the equation than it is to adjust the decay rate, although model accuracy does slightly improve if a higher decay rate is chosen with the standard form.
